---
title: "project1"
output: html_document
---
##Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")
```

##What is mean total number of steps taken per day?

First, we want to summarize the date using the following r code:
```{r}
library(dplyr)
group_activity <- group_by(activity, date)
steps <- summarize(group_activity, totalsteps = sum(steps))
```

Then, we made a histogram of the total number of steps taken each day.
```{r}
library(ggplot2)
qplot(steps$totalsteps, 
      geom="histogram",
      bins=8,
      main = "Histogram for Steps",
      xlab = "Total Steps per Day",
      ylab = "count",
      fill = I("blue"),
      color = I("black"),
      alpha = I(.5)
      )
```

Finally, we can calculate the mean and median total number of steps taken per day by the following r code:
```{r mean&median}
meansteps <- mean(steps$totalsteps, na.rm = TRUE)
mediansteps <- median(steps$totalsteps, na.rm = TRUE)
```

The mean total number of steps taken per day is `r meansteps`;
The median total number of steps taken per day is `r mediansteps`.

##What is the average daily activity pattern?

First, we summarized the mean steps in each 5-min interval over all days.
```{r}
group_activity2 <- group_by(activity, interval)
intervals <- summarize(group_activity2, msteps = mean(steps, na.rm=TRUE))
```

Then, a time series plot representing the mean steps was plotted.
```{r}
qplot(intervals$interval, intervals$msteps,
      geom="line",
      main = "Mean Steps over 5min Intervals",
      xlab = "Interval",
      ylab = " Mean Steps",
      color = I("blue")
      )
```

Further, the interval that contains the maximum value of mean steps can be found using the following r code:
```{r}
sortintervals <- arrange(intervals, desc(msteps))
maxinterval <- sortintervals[1,1]
```

and this interval is `r maxinterval`.

##Imputing missing values

First, here is the r code to calculate the total number of missing values in the dataset.
```{r}
missing <- sum(is.na(activity$steps))
```
There are `r missing` missing values(NA) in the dataset.

We chose to use the mean steps in the correspondent interval to replace the NA value.

Here is the r code to make the new dataset:

```{r}
nadata <- filter(activity, is.na(steps)==TRUE)
nonmissing <- filter(activity, is.na(steps)==FALSE)
joint <- merge(nadata, intervals, by="interval")
replacena <- joint %>% mutate(steps=msteps) %>% select(-msteps)
activityfilled <- rbind(replacena, nonmissing)
```

Now we make the histogram of the total number of steps each day and calculate the mean and median total number steps taken per day based on the new dataset.

```{r}
group_activityfilled <- group_by(activityfilled, date)
steps2 <- summarize(group_activityfilled, totalsteps = sum(steps))
meansteps2 <- mean(steps2$totalsteps, na.rm = TRUE)
mediansteps2 <- median(steps2$totalsteps, na.rm = TRUE)
```
```{r}
qplot(steps$totalsteps, 
      geom="histogram",
      bins=8,
      main = "Histogram for Steps",
      xlab = "Total Steps per Day",
      ylab = "count",
      fill = I("blue"),
      color = I("black"),
      alpha = I(.5)
      )
```

The mean and median total number steps taken per day are `r meansteps2` and `r mediansteps2`, respectively. After imputing NA by mean steps value of correspondent interval, there is no change for overall mean steps, but the median value changed.

##Are there differences in activity patterns between weekdays and weekends?

To answer this question, we first created a new column "weekday" in the new dataset using the following code:

```{r}
library(lubridate)
activitydate <- mutate(activityfilled, date=ymd(date))
activitydate <- mutate(activitydate, weekday=weekdays(date))
weekdays <- filter(activitydate, weekday %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
weekends <- filter(activitydate, weekday %in% c("Saturday", "Sunday"))
weekdays <-mutate(weekdays, weekday="weekday")
weekends <- mutate(weekends, weekday="weekend")
alldays <- rbind(weekends, weekdays)
alldays2 <- alldays %>% group_by(weekday, interval) %>% 
            summarise(msteps=mean(steps))
```

Then, a time series plot of mean steps of each interval is made for weekends and weekdays, respectively.

```{r}
g <- ggplot(alldays2, aes(x=interval, y=msteps))
g + geom_line(color="blue") + labs(y = "Number of steps") + facet_grid(weekday~.)
```

It seems there are indeed different patterns of mean steps between weekdays and weekends.